Implications of afforestation policies for global markets

### 2024-02-25 paper.Rmd
Update paper.Rmd

More recent modeling attempts of land-use change have begun to move away from comparative static analysis. For example, the dynamic GTAP model, supported by the GTAP-AEZ database, considers land market effects, which were identified as significant in driving results by Stevenson et al. [@stevenson2013], and has been utilized to evaluate the impacts of crop intensification on land use. It, however, has been critiqued, as global aggregates may mask localized shifts that can have implications for ecologically significant areas [@byerlee2014]. This project will attempt to get around these concerns and assess the impact of various likely afforestation scenarios by melding GTAP with the Global Timber Model (GTM), which is a dynamic optimization model that can examine the impact of afforestation or avoided deforestation on the carbon sink [@sohngen2001]. By joining these two models and passing outputs back and forth between the models, any limitations of either model can be minimized and more robust results, therefore, can be attained.

@@

The standard GTAP model is a leading CGE representation of the global economy and its trade linkages. Its general equilibrium nature means that all markets that are considered in the model are cleared with market-clearing prices. There are 68 market sectors covered in the latest version of the GTAP database includes  sectors. As the model applies nested CES production functions to all sectors, differentiating them by parameters only, it is difficult to model a sector individually. To model one sector, in this case forestry, differently from the rest would require a major redesign of the model. While not impossible, we chose to create a connection between the standard GTAP model and the GTM model, as GTM has been specifically designed to analyze the relationship among land rents, forested land cover, and carbon sequestration through forests. Doing so obviated the need to reparametrize the GTAP model for all countries.

>>>>

More recent modeling attempts of land-use change have begun to move away from comparative static analysis. For example, the dynamic GTAP model, supported by the GTAP-AEZ database, considers land market effects, which were identified as significant in driving results by Stevenson et al. [@stevenson2013], and has been utilized to evaluate the impacts of crop intensification on land use. It, however, has been critiqued, as global aggregates may mask localized shifts that can have implications for ecologically significant areas [@byerlee2014]. In addition, in the standard GTAP model, it is difficult to model a sector individually. There are 68 market sectors covered in the latest version of the GTAP database. The GTAP model's general equilibrium nature means that all these markets are cleared with market-clearing prices. As the model applies nested CES production functions to all sectors and differentiates them by parameters only, to model one sector, in this case forestry, differently from the rest would require a major redesign of the model. While not impossible, this project will attempt to get around these concerns and assess the impact of various likely land-use scenarios by melding GTAP with the Global Timber Model (GTM), which is a dynamic optimization model that has been specifically designed to analyze the relationship among land rents, forested land cover, and carbon sequestration through forests and thus is better able to examine the impact of afforestation or avoided deforestation on the carbon sink [@sohngen2001]. Joining these two models and passing outputs back and forth between the models should enable any limitations of either model to be minimized and more robust results, therefore, to be attained.

@@
